\documentclass[12pt]{report}
%%% Packages
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\usepackage{titlesec}
\usepackage{tocloft}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{array}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{graphicx}
\usepackage{subcaption}
%%%

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

%%%
\setlength{\parindent}{0pt}


\begin{document}

\title{Randomised Methods For Tensor-Product Preconditioners In Very High Order Discontinuous Galerkin Discretizations}
\author{Tom Feldhausen}
\date{\today}

\maketitle

\begin{abstract}
Solving a multi-dimensional PDE by means of a Discontinuous Galerkin method in a higher order solution space requires the solution of large linear systems during time integration. When storing the semi-discrete system, communication dominates the computational costs such that matrix-free implementations~\cite{kronbichler2019fast} often turn out to be more efficient on modern architectures. In order to efficiently integrate implicit time integration into this framework, we need matrix-free tensor-product preconditioners such as the tensor-product preconditioners of \cite{diosady2017tensor} and \cite{pazner2018approximate}. In the first part of the project, we implemented a matrix-based DG method for two-dimensional linear advection with upwind flux and verified the efficiency of Pazner's preconditioner. Next, we aim to reduce the number of iterations by enhancing the preconditioner with randomised methods.
\end{abstract}

\section{Introduction To Matrix-Free Tensor-Product Preconditioning}
In the first part of a project, we derived and implemented a 2D Discontinuous Galerkin scheme for the linear advection equation, giving rise to an element-wise system of the form
\begin{align}
\frac{dU(t)}{dt} = M^{-1} [(B - G) U - G_{\textnormal{bound}}].
\end{align}

When discretized with hexahedral elements, the involved matrices have tensor-product form. Specifically, let $x_i$ denote quadrature points, $w_i$ quadrature weights, $\phi_j$ 1D basis functions and $T$ the reference mapping. We introduce the 1D operators
\begin{align}
&G_{i,j} = \phi_j(x_i), \quad D_{i,j} = D \phi_j(x_i)
\\
&W_{i, i} = w_i, \quad E_{L,j} = \phi_j(-1), \quad E_{R, j} = \phi_j(1) 
\\
&J_T = \det(T(\cdot)), \quad \Lambda_{i,i} = \hat{au}(x_i).
\end{align}

With this notation at hand, we can express the DG matrices in cases with constant velocity 1 as
\begin{align}
&M = ((G^\top W) \otimes (G^\top W)  \otimes (G^\top W)) J_T (G \otimes ... \otimes G)
\\
&B = ((G^\top W) \otimes (G^\top W)  \otimes (G^\top W)) [(D \otimes G \otimes G) + (G \otimes D \otimes G) + (G \otimes G \otimes D)]
\\
&G_{\textnormal{2D face}} = (E_R \otimes G \otimes G)^\top (I \otimes W \otimes W) \Lambda (E_R \otimes G \otimes G).
\end{align}

For the time discretization, we compared explicit and SDIRK implicit schemes. The latter give rise to element-wise linear systems of the form $A_e x = b$ with
\begin{align}
A_e = M_e - h a_{j,j} (G - B).
\end{align}

We applied the tensor product preconditioners of \citep{diosady2017tensor} and \cite{pazner2018approximate}, which exhibit a large reduction in the number of iterations of the iterative linear solvers, i.e. block-Krylov methods. \\

Traditionally, ILU or Gauss--Seidel preconditioners were used, but they require more than $\mathcal{O}(p^{d+1})$, which is the efficiency of matrix-free vector evaluations of the linear operator. \\

Diosady~\citep{diosady2017tensor} suggest two approaches that exploits the tensor-product structure of the operator, the first one being the ADI scheme. \cite{diosady2017tensor} reduces the linear system to a matrix of the form 
\begin{align*}
.
\end{align*}

Running one iteration of alternating least squares, we arrive at the approximate inverse
\begin{align*}
.
\end{align*}

While this approach is very fast, it is only a rough approximation that does not suffice to precondition complex behaviour well. \\

Alternatively, Diosady~\cite{diosady2017tensor} suggests the FDM preconditioner. If we can write $D_i = X_i \Lambda_i X_i^{-1}$, the inversion is easily determined as 
\begin{align*}
.
\end{align*}

Once again we benefit from high efficiency at the expense of ill-conditioning of $X_i$ and strong structural assumptions. While the approach works naturally for Laplacians that can be written as $\Delta_{x,y} = D_x \otimes I + I \otimes D_y$, Navier--Stokes equations rely on cross-derivatives such as $D_x \otimes D_y \otimes I$. \\

Here enters Pazner's~\cite{pazner2018approximate} preconditioner. Using a Lanczos SVD on a reshuffled operator, we can compute 
\begin{align}
A_e \approx \sum_{k=1}^r A_k \otimes B_k \otimes C_k
\end{align}
in $\mathcal{O}(p^{d+1})$ time. For $r = 1$, the inversion is straightforward and for $r = 2$, it is efficiently found solving a Sylvester system. However, $r \leq 2$ might be inaccurate. While KSVD rank might be low, it still induces a full-rank matrix. This is crucial because the dynamics of fluid flow are full-rank, whereas the turbulences are low-rank phenomenon.

\section{Strategies For Randomized Matrix-Free Tensor-Product Preconditioning}
We present the following preliminary ideas to enhance preconditioning via randomized means. We quickly summarize:
\begin{enumerate}
\item \textbf{Blendenpik}: Our source of inspiration, unsuited for our problem.

\item \textbf{Randomized GMRES}: Randomized iterative solvers, but not preconditioners.

\item \textbf{RSVD for KSVD}: State-of-the-art implementation for Pazner,

\item \textbf{Tensor Formats for higher ranks}: The KSVD is cheaply transformed to other tensor formats, which allow cheap application of the inverse.

\item \textbf{Woodbury}: Combine cheap full-rank base preconditioner (ADI, FDM, Pazner) for the fluid flow with low-rank preconditioner for turbulences.
\end{enumerate}

\subsubsection{Blendenpik}
While unhelpful for our means, we draw inspiration from the Blendenpik algorithm. Choosing $R$ from $\Psi^\top A_e = QR$ as a preconditioner, it was one of the first randomized algorithms outperforming deterministic counterparts for least squares approximation. Problematically however, this ansatz does not address the full-rank nature of fluid dynamics nor the tensor-product structure of the problem.

\subsubsection{Randomized GMRES}
Suggested by Nakatsukasa and Tropp~\cite{}, the randomized variant promises significant runtime advantages at near-optimal accuracy by changing the iterative solver.

\subsubsection{Efficient KSVD}
The Lanczos SVD is no longer state of the art for small-rank truncated SVDs, but the randomized SVD~\cite{} is, such that we should include this algorithm. For rank $r$, it requires $2r$ matrix-vector products, Lanczos requires $2J$, where $J \geq r$ is the number of iterations.

\subsubsection{The Inversion Of Higher Ranks}
The KSVD can be cheaply transformed to a formulation similar to the Tucker format. We derive it in the following but realize it suffers from the curse of dimensionality. \\

Start with $E = \sum_{k=1}^r A_k \otimes B_k \otimes C_k$. $E$ is $N \times N$ with $N = p^d$ and $A_k, B_k, C_k$ are $p \times p$. We sketch cheaply with $\Omega = \Omega_x \otimes \Omega_y \otimes \Omega_z$ with $\Omega$ $N \times k^3$ and $\Omega_x, \Omega_y, \Omega_z$ $p \times k$ such that in $\mathcal{O}(dr k p^2)$ time we arrive at the factorization
\begin{align}
E \Omega = \sum_{k=1}^r (A_k \Omega_x) \otimes (B_k \Omega_y) \otimes (C_k \Omega_z).
\end{align}

Our bases are determined by computing
\begin{align*}
&Q_x = \begin{pmatrix}A_1 \Omega_x & ... & A_r \Omega_x \end{pmatrix}
\\
&Q_y = \begin{pmatrix}B_1 \Omega_y & ... & B_r \Omega_y \end{pmatrix}
\\
&Q_z = \begin{pmatrix}C_1 \Omega_z & ... & C_r \Omega_z \end{pmatrix}
\end{align*}
in $\mathcal{O}(p (kr)^2)$ time for each basis with all bases of shape $p \times (rk)$. \\

We can finally compute the factorization of the $k^d \times k^d$ core in $\mathcal{O}(2rd p (rk)^2)$ as 
\begin{align*}
C = (Q_x^\top \otimes Q_y^\top \otimes Q_z^\top) E (Q_x \otimes Q_y \otimes Q_z) = \sum_{k=1}^r (Q_x^\top A_k Q_x) \otimes (Q_y^\top B_k Q_y) \otimes (Q_z^\top C_k Q_z).
\end{align*}

Once we have our Tucker format $E^{-1} \approx (Q_x^\top \otimes Q_y^\top \otimes Q_z^\top) C^{-1} (Q_x \otimes Q_y \otimes Q_z)$ is applied in $\mathcal{O}(p^{d+1} + k^{3d})$ time, which is too expensive. For higher ranks, inversion seems to remain a problem.

\subsubsection{Woodbury}
Naturally, $A_e = P_1 + E$ can be decomposed by choosing the base preconditioner $P_1$ as Pazner / FDM / ADI. We can apply $E$ in $\mathcal{O}(p^{d+1})$ time. Writing $P \approx P_1 + U V^\top$ with $U, V$ $N \times r$, Woodbury's formula~\cite{} gives 
\begin{align*}
z = P^{-1} v \approx P_1^{-1} v - P_1^{-1} U (I_k + V^\top P_1^{-1} U)^{-1} V^\top P_1^{-1} v.
\end{align*}

Computing the factorization $E \approx UV^\top$ costs $\mathcal{O}(4k d p^{d+1})$ using the randomized SVD. \\

For the evaluation, see that applying the base preconditioner $w = P_1^{-1} v$ takes $\mathcal{O}(d p^{d+1})$ time and calculating the dense product $x = V^\top w$ takes $\mathcal{O}(k p^d)$ time. The core solve $(I_k + V^\top P_1^{-1} U)^{-1} x$ takes $\mathcal{O}(k^2)$ time, where the precomputation of the core costs $\mathcal{O}(k d p^{d+1})$. Applying $P_1^{-1}U$ (forming it is part of the prior precomputation) costs $\mathcal{O}(k^2 p^d)$ time. We conclude that the extra costs of the Woodbury update is $\mathcal{O}((4k + 2k + 1)d p^{d+1})$ for precomputation, whereas the update is free in the sense that it is lower-order. We expect massive computational advantages.

\section{Randomized Woodbury Preconditioning}
We implement the Woodbury approach and apply it to the exact same problem as in the first part of the project, linear advection with constant velocity. It gets the same iteration numbers as Pazner's preconditioner. This means two things: it works and the problem is too simple such that Pazner models the underlying geometry remarkably well. It is not exact because we move in a block-Jacobi setting, but it is exact per block. \\

This is why we increase the complexity of the problem. Instead of using a constant velocity, we use a rotating vortex velocity field $v(x, y) = (-y, x)$. For $r = 2$ we have the same picture due to the fact that the continuous operator is a rank-$2$ Kronecker product, i.e.
\begin{align}
-y \partial_x u + x \partial_y u = D_x \otimes M_y(-y) + M_x(x) D_y.
\end{align}

We expect this to no longer be a case in 3D. Alternatively, we can use a worse preconditioner. Either Pazner with $r = 1$ or simply the matrix diagonal. The latter might be too bad to capture the full-rank fluid dynamics and thus requires massive rank in our Woodbury improvement, which is likely too costly. The first option is thus preferred and the results are promising. The higher the rank of our Woodbury patch, the lower the number of iterations. Ranks $k = 8-16$ even deliver as few iterations as the $r = 2$ Pazner preconditioner. This is remarkable because it suggests that in 3D, combining $r = 2$ and moderately high $k$ can reduce the number of iterations to $r = 3$ Pazner preconditioning, which is infeasible in practice without losing $\mathcal{O}(p^{d+1})$. \\

\begin{table}[h!]
\centering
\begin{tabular}{c|cccc|cccc}
\hline
 & \multicolumn{4}{c|}{\textbf{Pazner ($r=1$)}} 
 & \multicolumn{4}{c}{\textbf{Pazner ($r=2$)}} \\
\textbf{RK Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
1 & 61.00 & 38.00 & 15.20 & 11.75 
  & 19.00 & 15.00 & 9.60 & 8.10 \\
2 & 53.00 & 33.50 & 14.00 & 10.78 
  & 18.50 & 13.00 & 9.00 & 8.00 \\
3 & 64.67 & 40.00 & 15.93 & 12.00 
  & 20.00 & 16.00 & 10.00 & 8.45 \\
4 & 27.75 & 17.25 & 8.25 & 6.68 
  & 10.50 & 8.25 & 6.00 & 5.74 \\
\hline
\end{tabular}
\caption{Average number of GMRES iterations for RK methods using Pazner's preconditioner with rank $r=1$ versus rank $r=2$ with respect to step sizes $h$ and stage numbers.}
\label{tab:pazner_comparison}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{c|cccc|cccc}
\hline
 & \multicolumn{4}{c|}{\textbf{Diagonal Preconditioner}} 
 & \multicolumn{4}{c}{\textbf{Without Preconditioner}} \\
\textbf{RK Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
1 & 150.00 & 150.00 & 39.00 & 23.00 
  & 150.00 & 150.00 & 29.30 & 14.65 \\
2 & 150.00 & 136.25 & 32.40 & 19.00 
  & 150.00 & 136.75 & 23.70 & 12.43 \\
3 & 150.00 & 150.00 & 41.70 & 24.00 
  & 150.00 & 150.00 & 32.50 & 15.65 \\
4 & 108.25 & 59.00  & 15.98 & 9.75  
  & 112.50 & 55.13  & 9.95  & 6.58  \\
\hline
\end{tabular}
\caption{Average number of GMRES iterations for RK methods using the Diagonal preconditioner compared to a plain implicit solver without preconditioning. Notably, for smaller step sizes ($h \le 0.01$), the diagonal preconditioner degrades the spectrum and increases iteration counts compared to the un-preconditioned system. Values of 150.00 indicate GMRES stalled at the maximum iteration limit.}
\label{tab:diagonal_vs_none}
\end{table}

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{c|cccc|cccc|cccc|cccc}
\hline
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=2$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=4$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=8$}} 
 & \multicolumn{4}{c}{\textbf{Woodbury $k=16$}} \\
\textbf{RK Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
1 & 55.00 & 35.50 & 15.00 & 12.00 
  & 44.00 & 31.50 & 14.00 & 11.00 
  & 30.00 & 22.00 & 12.00 & 10.00 
  & 19.00 & 15.00 & 9.60  & 8.00 \\
2 & 45.00 & 31.00 & 14.00 & 11.00 
  & 38.00 & 27.50 & 13.00 & 10.43 
  & 27.00 & 20.25 & 11.20 & 9.00 
  & 18.50 & 13.00 & 9.00  & 8.00 \\
3 & 55.33 & 36.00 & 15.87 & 12.00 
  & 49.67 & 32.17 & 15.00 & 11.17 
  & 32.67 & 22.00 & 12.00 & 10.00 
  & 19.00 & 16.00 & 10.00 & 8.27 \\
4 & 24.00 & 16.50 & 8.25  & 6.75 
  & 21.00 & 15.75 & 8.25  & 6.75 
  & 16.50 & 11.38 & 7.50  & 6.00 
  & 10.50 & 8.25  & 6.00  & 5.56 \\
\hline
\end{tabular}
}
\caption{Average number of GMRES iterations for RK methods using Pazner's $r=1$ base preconditioner augmented with a randomized Woodbury update of sketch ranks $k \in \{2, 4, 8, 16\}$. Notice the convergence to the exact $r=2$ iteration baseline at $k=16$.}
\label{tab:woodbury_k_comparison}
\end{table}

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{c|cccc|cccc|cccc|cccc}
\hline
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=2$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=4$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=8$}} 
 & \multicolumn{4}{c}{\textbf{Woodbury $k=16$}} \\
\textbf{RK Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
1 & 2.58 & 3.00 & 7.81  & 13.70 
  & 2.81 & 3.57 & 8.98  & 15.92 
  & 2.42 & 3.06 & 9.04  & 16.27 
  & 1.98 & 2.25 & 7.07  & 12.33 \\
2 & 3.44 & 4.94 & 15.19 & 31.54 
  & 3.55 & 5.27 & 16.12 & 30.50 
  & 3.28 & 4.82 & 15.30 & 27.84 
  & 2.41 & 3.58 & 12.64 & 23.78 \\
3 & 7.42 & 10.85 & 28.00 & 44.13 
  & 5.88 & 9.20 & 26.39 & 45.14 
  & 4.47 & 6.77 & 22.72 & 40.45 
  & 3.25 & 5.37 & 19.67 & 35.93 \\
4 & 4.29 & 6.77 & 21.83 & 38.48 
  & 4.21 & 6.30 & 20.01 & 35.01 
  & 3.15 & 4.64 & 15.68 & 28.17 
  & 2.73 & 4.05 & 14.61 & 28.07 \\
\hline
\end{tabular}
}
\caption{Total runtime (in seconds) for RK methods using Pazner's $r=1$ base preconditioner augmented with a randomized Woodbury update of sketch ranks $k$. Notice that increasing the rank to $k=16$ yields the fastest overall runtimes due to the drastic reduction in GMRES iterations.}
\label{tab:woodbury_runtime}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{c|cccc|cccc}
\hline
 & \multicolumn{4}{c|}{\textbf{Pazner ($r=1$)}} 
 & \multicolumn{4}{c}{\textbf{Pazner ($r=2$)}} \\
\textbf{RK Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
1 & 2.32 & 3.94 & 7.66  & 13.11 
  & 2.14 & 2.61 & 8.31  & 14.95 \\
2 & 3.48 & 4.80 & 13.73 & 34.26 
  & 2.90 & 4.14 & 15.57 & 28.97 \\
3 & 5.66 & 7.47 & 21.79 & 39.04 
  & 4.25 & 6.91 & 24.82 & 45.31 \\
4 & 3.75 & 5.33 & 15.37 & 28.08 
  & 3.28 & 5.15 & 19.31 & 36.47 \\
\hline
\end{tabular}
\caption{Total runtime (in seconds) for RK methods using Pazner's base preconditioner with rank $r=1$ versus rank $r=2$.}
\label{tab:pazner_runtime_comparison}
\end{table}

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{c|cccc|cccc|cccc|cccc}
\hline
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=2$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=4$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=8$}} 
 & \multicolumn{4}{c}{\textbf{Woodbury $k=16$}} \\
\textbf{RK Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
1 & 150.00 & 145.50 & 33.70 & 21.00 
  & 150.00 & 133.00 & 29.00 & 19.00 
  & 149.00 & 75.50  & 20.00 & 15.00 
  & 19.00  & 15.00  & 9.60  & 8.00  \\
2 & 150.00 & 107.00 & 29.00 & 19.03 
  & 150.00 & 92.50  & 24.65 & 17.00 
  & 115.50 & 51.00  & 18.10 & 13.00 
  & 18.50  & 13.00  & 9.00  & 8.00  \\
3 & 150.00 & 150.00 & 35.03 & 21.78 
  & 150.00 & 137.67 & 30.00 & 19.17 
  & 138.67 & 69.83  & 21.07 & 15.00 
  & 19.00  & 16.00  & 10.00 & 8.27  \\
4 & 89.25  & 45.38  & 14.93 & 10.46 
  & 88.25  & 39.00  & 12.75 & 9.45  
  & 48.00  & 25.50  & 10.50 & 8.25  
  & 10.50  & 8.25   & 6.00  & 5.56  \\
\hline
\end{tabular}
}
\caption{Average number of GMRES iterations for RK methods using a Diagonal base preconditioner augmented with a randomized Woodbury update. Notice that at $k=16$, the randomized update completely maps the missing cross-derivative physics, rescuing the solver from stagnation and achieving optimal iteration counts.}
\label{tab:diagonal_woodbury_comparison}
\end{table}

\textbf{Enter nice graphics and present to Martin.}

Next, we can do the following:
\begin{enumerate}
\item Run for higher-order basis functions.
\item Add more complicated initial value, see advection miniapp.
\item Consider different base preconditioners: FDM, ADI, matrix diagonal.
\item Extend to 3D.
\item Optimize code in deal.II.
\item Add diffusion terms, Navier--Stokes.
\end{enumerate}

We do the same experiment for constant advection. Maximum number of iterations is 150. Pazner1 with Woodbury is the basis. Woodbury does not change the results of Pazner2 (up to minor pertubrations +-2). We choose Nx = Ny = 8. We choose J = 10 for the Lanczos iteration, but note that forming the KSVD has the same runtime for all methods considered (independent of J). However, application of the r = 2 Pazner method is much more expensive than for r = 1. We choose final time T = 0.1 because my PC is slow.

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{cc|cccc|cccc|cccc|cccc|cccc|cccc}
\hline
& & \multicolumn{4}{c|}{\textbf{Woodbury $k=0$}}
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=2$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=4$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=8$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=16$}} 
 & \multicolumn{4}{c}{\textbf{Pazner $r=2$}} \\
\textbf{Deg.} & \textbf{Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
\hline
\textbf{$p=3$} 
& 1 & 77.00 & 48.00 & 17.00 & 12.00 & 86.00 & 49.00 & 18.00 & 13.60 & 75.00 & 43.00 & 17.00 & 13.00 & 48.00 & 30.00 & 14.80 & 12.00 & 16.00 & 17.00 & 9.00  & 8.00 & 16.00 & 17.00 & 9.00  & 8.00 \\
& 2 & 65.50 & 41.00 & 16.50 & 11.00 & 71.00 & 41.00 & 16.00 & 12.00 & 62.00 & 36.00 & 16.00 & 12.00 & 40.00 & 25.00 & 13.00 & 11.00 & 16.00 & 16.00 & 10.00 & 8.00 & 16.00 & 16.00 & 10.00 & 8.00 \\
& 3 & 78.67 & 51.00 & 18.00 & 13.00 & 94.00 & 51.00 & 18.00 & 14.00 & 79.00 & 45.00 & 19.00 & 13.00 & 52.67 & 31.00 & 15.00 & 11.68 & 16.00 & 15.00 & 9.00  & 8.00 & 16.00 & 15.00 & 9.00  & 8.00 \\
& 4 & 33.00 & 21.00 & 9.00  & 6.75  & 32.25 & 20.25 & 9.75  & 7.50  & 29.25 & 18.75 & 9.30  & 7.50  & 20.25 & 14.63 & 8.25  & 6.75  & 12.00 & 8.25  & 6.00  & 5.25 & 12.00 & 8.25  & 6.00  & 5.25 \\
\hline
\textbf{$p=4$} 
& 1 & 129.00 & 65.00 & 23.00 & 15.00 & 125.00 & 67.00 & 22.00 & 15.00 & 112.00 & 65.00 & 20.00 & 15.00 & 92.00 & 50.00 & 18.80 & 14.00 & 45.00 & 28.50 & 13.00 & 10.00 & 21.00 & 16.00 & 9.00 & 8.00 \\
& 2 & 102.00 & 54.00 & 18.00 & 13.00 & 101.00 & 54.00 & 18.00 & 14.00 & 92.00 & 50.00 & 19.00 & 14.73 & 77.00 & 42.00 & 18.00 & 13.00 & 38.00 & 24.00 & 12.00 & 10.00 & 20.00 & 15.00 & 9.00 & 8.00 \\
& 3 & 128.00 & 70.00 & 22.00 & 15.00 & 129.67 & 70.00 & 21.97 & 17.00 & 118.67 & 65.00 & 22.00 & 15.00 & 103.00 & 53.00 & 19.00 & 14.00 & 50.00 & 29.00 & 13.00 & 11.00 & 20.00 & 18.00 & 9.00 & 8.00 \\
& 4 & 43.50 & 24.75 & 10.50 & 8.25  & 44.25 & 25.50 & 10.50 & 8.25  & 41.25 & 24.00 & 10.50 & 8.25  & 33.75 & 21.00 & 10.50 & 7.50  & 18.75 & 13.50 & 7.50  & 6.00  & 12.00 & 9.00 & 6.00 & 5.25 \\
\hline
\textbf{$p=5$} 
& 1 & 150.00 & 87.00 & 24.00 & 16.00 & 150.00 & 85.00 & 24.00 & 18.00 & 150.00 & 80.50 & 23.00 & 18.00 & 140.00 & 73.00 & 22.00 & 16.00 & 97.00 & 53.00 & 19.40 & 14.00 & 24.00 & 17.00 & 9.00 & 8.00 \\
& 2 & 126.00 & 67.00 & 22.00 & 16.00 & 130.00 & 69.00 & 22.00 & 15.00 & 124.00 & 65.00 & 20.00 & 15.00 & 109.00 & 60.00 & 19.00 & 15.00 & 76.00 & 45.00 & 17.00 & 13.00 & 23.00 & 16.00 & 9.00 & 8.00 \\
& 3 & 150.00 & 92.83 & 25.00 & 17.00 & 150.00 & 91.00 & 24.40 & 17.00 & 150.00 & 86.00 & 24.00 & 17.00 & 147.33 & 76.00 & 23.00 & 16.00 & 106.00 & 57.00 & 21.00 & 15.00 & 25.33 & 17.00 & 11.00 & 8.00 \\
& 4 & 57.75 & 30.00 & 11.25 & 9.75  & 57.00 & 30.75 & 12.00 & 9.00  & 54.00 & 30.00 & 12.75 & 9.00  & 48.75 & 27.75 & 11.25 & 9.00  & 36.00 & 22.50 & 10.50 & 8.25  & 12.00 & 9.00 & 6.00 & 5.25 \\
\hline
\textbf{$p=6$} 
& 1 & 150.00 & 102.00 & 26.00 & 20.00 & 150.00 & 105.00 & 27.00 & 20.00 & 150.00 & 102.00 & 27.00 & 18.00 & 150.00 & 96.00 & 26.00 & 19.00 & 150.00 & 79.00 & 23.00 & 16.00 & 24.00 & 17.00 & 10.00 & 8.00 \\
& 2 & 150.00 & 84.50 & 23.00 & 18.00 & 150.00 & 84.00 & 23.00 & 18.00 & 150.00 & 82.00 & 23.00 & 16.00 & 148.00 & 77.00 & 22.00 & 16.00 & 121.00 & 65.00 & 20.00 & 15.00 & 21.00 & 16.00 & 9.00 & 8.00 \\
& 3 & 150.00 & 112.33 & 30.00 & 20.00 & 150.00 & 114.17 & 28.00 & 19.00 & 150.00 & 108.00 & 28.00 & 20.00 & 150.00 & 102.00 & 27.00 & 18.00 & 150.00 & 84.00 & 24.00 & 17.00 & 24.00 & 18.00 & 10.00 & 9.00 \\
& 4 & 71.25 & 36.75 & 12.75 & 9.75  & 69.75 & 37.50 & 12.75 & 9.75  & 67.50 & 36.75 & 12.75 & 9.75  & 63.00 & 34.50 & 13.50 & 9.75  & 53.25 & 30.88 & 11.25 & 9.00  & 11.25 & 9.75 & 6.00 & 5.25 \\
\hline
\end{tabular}
}
\caption{Average GMRES iterations per RK step for varying polynomial degrees $p$. The values indicate the convergence behavior of the randomized Woodbury update across different ranks $k$ and time steps.}
\label{tab:stacked_woodbury_iterations}
\end{table}

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{cc|cccc|cccc|cccc|cccc|cccc|cccc}
\hline
& & \multicolumn{4}{c|}{\textbf{Woodbury $k=0$}}
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=2$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=4$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=8$}} 
 & \multicolumn{4}{c|}{\textbf{Woodbury $k=16$}} 
 & \multicolumn{4}{c}{\textbf{Pazner $r=2$}} \\
\textbf{Deg.} & \textbf{Stages} 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 
& 0.1 & 0.05 & 0.01 & 0.005 \\
\hline
\hline
\textbf{$p=3$} 
& 1 & 1.09 & 1.06 & 1.84 & 2.26 & 1.36 & 1.74 & 2.88 & 3.06 & 0.83 & 1.17 & 2.22 & 2.84 & 0.69 & 1.16 & 1.81 & 2.99 & 0.58 & 0.98 & 1.67 & 2.43 & 0.64 & 0.99 & 2.06 & 3.42 \\
& 2 & 0.92 & 1.40 & 2.67 & 4.16 & 1.36 & 1.92 & 3.64 & 4.74 & 1.39 & 1.99 & 2.89 & 4.72 & 1.48 & 1.65 & 2.87 & 5.89 & 1.00 & 1.06 & 3.30 & 4.34 & 0.80 & 1.34 & 4.12 & 6.87 \\
& 3 & 1.41 & 2.02 & 4.60 & 6.39 & 2.10 & 2.50 & 5.73 & 7.05 & 1.95 & 2.42 & 7.70 & 8.77 & 1.95 & 2.18 & 3.93 & 7.27 & 0.81 & 1.17 & 3.33 & 6.71 & 1.18 & 1.74 & 5.53 & 10.30 \\
& 4 & 1.28 & 1.52 & 2.88 & 4.49 & 1.01 & 1.44 & 3.04 & 5.15 & 1.28 & 1.70 & 3.38 & 5.28 & 0.93 & 1.52 & 2.92 & 4.95 & 0.76 & 0.94 & 2.60 & 4.48 & 1.08 & 1.49 & 4.79 & 8.05 \\
\hline
\textbf{$p=4$} 
& 1 & 1.26 & 1.61 & 3.56 & 6.87 & 1.70 & 1.82 & 3.66 & 5.98 & 1.81 & 2.03 & 4.29 & 7.08 & 1.62 & 1.78 & 4.33 & 7.07 & 1.20 & 1.74 & 4.03 & 5.20 & 1.21 & 1.63 & 3.53 & 6.07 \\
& 2 & 2.45 & 2.98 & 7.54 & 12.28 & 2.07 & 2.51 & 6.29 & 11.23 & 2.52 & 3.22 & 7.01 & 12.08 & 1.98 & 2.35 & 6.91 & 16.03 & 1.59 & 2.04 & 6.12 & 8.70 & 1.70 & 2.43 & 6.41 & 11.31 \\
& 3 & 3.13 & 4.00 & 8.90 & 15.77 & 3.25 & 4.22 & 11.42 & 20.27 & 4.01 & 4.26 & 11.58 & 19.15 & 3.52 & 4.26 & 11.46 & 17.46 & 1.99 & 2.83 & 8.82 & 14.70 & 2.09 & 3.48 & 8.65 & 15.92 \\
& 4 & 1.76 & 2.24 & 6.29 & 11.97 & 2.44 & 3.00 & 7.98 & 14.79 & 2.24 & 2.72 & 6.85 & 12.39 & 1.83 & 2.36 & 7.27 & 9.13 & 1.58 & 2.13 & 5.09 & 8.67 & 1.77 & 2.72 & 6.86 & 12.13 \\
\hline
\textbf{$p=5$} 
& 1 & 2.37 & 3.13 & 6.23 & 10.56 & 2.86 & 3.04 & 6.08 & 10.50 & 2.90 & 2.99 & 6.17 & 10.77 & 3.02 & 2.92 & 6.23 & 10.38 & 2.93 & 2.86 & 6.07 & 10.18 & 2.45 & 3.11 & 6.91 & 11.74 \\
& 2 & 3.36 & 4.53 & 10.76 & 19.24 & 3.57 & 4.31 & 10.82 & 18.90 & 3.50 & 4.22 & 10.56 & 22.22 & 3.33 & 4.12 & 10.57 & 20.07 & 3.03 & 3.88 & 10.48 & 19.10 & 2.73 & 3.86 & 12.14 & 23.52 \\
& 3 & 4.31 & 5.69 & 15.88 & 30.11 & 6.21 & 7.42 & 15.83 & 28.33 & 5.18 & 7.05 & 17.69 & 30.08 & 5.28 & 6.08 & 15.72 & 28.08 & 4.39 & 5.55 & 18.49 & 30.63 & 4.25 & 5.29 & 18.86 & 34.91 \\
& 4 & 2.93 & 3.70 & 10.25 & 24.25 & 3.40 & 4.22 & 11.18 & 19.77 & 3.45 & 4.27 & 11.93 & 20.69 & 3.11 & 4.01 & 11.23 & 20.12 & 3.57 & 4.56 & 12.83 & 23.10 & 3.15 & 4.42 & 13.26 & 18.98 \\
\hline
\textbf{$p=6$} 
& 1 & 3.22 & 4.75 & 9.75 & 16.46 & 4.24 & 5.22 & 9.95 & 17.00 & 4.28 & 5.12 & 10.06 & 16.64 & 4.27 & 4.93 & 9.91 & 16.58 & 4.23 & 4.61 & 9.94 & 16.64 & 3.62 & 4.10 & 9.98 & 16.76 \\
& 2 & 5.15 & 6.52 & 16.63 & 36.82 & 5.52 & 6.98 & 17.07 & 30.30 & 5.56 & 6.87 & 17.02 & 30.33 & 5.36 & 6.60 & 16.94 & 30.84 & 5.27 & 6.72 & 17.04 & 30.90 & 4.19 & 5.66 & 17.10 & 31.30 \\
& 3 & 8.04 & 12.87 & 35.22 & 45.31 & 7.13 & 10.01 & 25.19 & 45.40 & 7.01 & 9.89 & 25.25 & 45.54 & 9.57 & 10.68 & 26.71 & 47.17 & 8.10 & 10.88 & 25.13 & 49.93 & 5.07 & 7.45 & 24.97 & 48.69 \\
& 4 & 5.03 & 6.34 & 17.39 & 32.23 & 5.29 & 6.64 & 17.63 & 33.00 & 5.28 & 6.58 & 17.52 & 31.75 & 6.74 & 7.81 & 19.67 & 31.83 & 6.06 & 7.66 & 19.10 & 34.36 & 4.60 & 6.00 & 18.36 & 34.04 \\
\hline
\end{tabular}
}
\caption{Total runtime (in seconds) per RK step for varying polynomial degrees $p$. The values illustrate the computational efficiency of randomized Woodbury updates as local cell matrix dimensionality increases.}
\label{tab:stacked_woodbury_runtimes}
\end{table}

%%% Bibliography
\bibliographystyle{alpha}
\bibliography{references}
%%%

\end{document}
